# Reproducible Research: Peer Assessment 1
#

## Loading and preprocessing the data
```{r Load data, echo = TRUE}
unzip("activity.zip")
data <- read.csv("activity.csv", colClasses = c("numeric","character","numeric"), na.strings = c("NA"))
```

## What is mean total number of steps taken per day?
```{r}
total_steps <- tapply(data$steps, data$date, sum)
hist(total_steps,ylim=c(1,30), main = "Frequency of steps taken per day", xlab = "Total steps per day")
```

#### Mean of total number of steps taken per day
```{r, results="asis"}
mean(total_steps,na.rm=TRUE)
```

#### Median of total number of steps taken per day
```{r, results="asis"}
median(total_steps,na.rm=TRUE)
```

## What is the average daily activity pattern?
```{r}
average_activity <- tapply(data$steps, data$interval, mean, na.rm = TRUE)
average_activity_df <- data.frame(key= as.numeric(names(average_activity)), value=as.numeric(average_activity))
colnames(average_activity_df) <- c("Interval", "AverageSteps")
plot(average_activity_df$AverageSteps~average_activity_df$Interval, type="l",ylab="Average Daily Activity", xlab ="Interval")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
average_activity_df[which.max(average_activity_df$AverageSteps),]$Interval
```

## Imputing missing values
### Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(data))
```

### Devise a strategy for filling in all of the missing values in the dataset.
#### If value in data is "NA", then replace with the average daily activity for that interval
```{r}
data_imputed <- within(data, steps <- ifelse(is.na(steps),average_activity_df[(average_activity_df$Interval==interval),]$AverageSteps,steps))
total_steps_imputed <- tapply(data_imputed$steps, data_imputed$date, sum)
hist(total_steps_imputed,ylim=c(1,30), main = "Frequency of steps taken per day (Imputed Data)", xlab = "Total steps per day")
```

#### Mean of total number of steps taken per day (Imputed)
```{r, results="asis"}
mean(total_steps_imputed,na.rm=TRUE)
```

#### Median of total number of steps taken per day (Imputed)
```{r, results="asis"}
median(total_steps_imputed,na.rm=TRUE)
```

#### As we can see from the values above the imputed data histogram, mean and median produce the same results as the original, non-imputed, data

## Are there differences in activity patterns between weekdays and weekends?
### Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
data_imputed$date <- strptime(data_imputed$date, format='%Y-%m-%d')
data_imputed$Day <-  ifelse(weekdays(data_imputed$date) %in% c("Saturday", "Sunday"), "Weekend","Weekday")
data_imputed$Day <- as.factor(data_imputed$Day)
```

### Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
data_weekday = subset(data_imputed, Day=="Weekday")
average_activity_weekday <- tapply(data_weekday$steps, data_weekday$interval, mean, na.rm = TRUE)
average_activity_weekday_df <- data.frame(key= as.numeric(names(average_activity_weekday)), value=as.numeric(average_activity_weekday))
colnames(average_activity_weekday_df) <- c("Interval", "AverageSteps")

data_weekend = subset(data_imputed, Day=="Weekend")
average_activity_weekend <- tapply(data_weekend$steps, data_weekend$interval, mean, na.rm = TRUE)
average_activity_weekend_df <- data.frame(key= as.numeric(names(average_activity_weekend)), value=as.numeric(average_activity_weekend))
colnames(average_activity_weekend_df) <- c("Interval", "AverageSteps")

par(mfrow = c(2,1))
plot(average_activity_weekend_df$AverageSteps~average_activity_weekend_df$Interval, type="l",ylab="Average Weekend Activity", xlab ="Interval")
plot(average_activity_weekday_df$AverageSteps~average_activity_weekday_df$Interval, type="l",ylab="Average Weekday Activity", xlab ="Interval")
```


